Vragenlijst thesis evaluatie

### 2015-08-18 doc/template/chap-9.tex
shizzle

\begin{figure}[hbtp]
\centering
\includegraphics[scale=1]{Images/functie4.pdf} 
\caption{The benefit function of the defender for a cost $k_{D} =1$, $d =2$ and $\delta_{D}=1.5$. The three pieces match with the three colours.}
\label{pfb}
\end{figure}

The benefit formula for the attacker:
\begin{displaymath}
\beta_{A}(\delta_{D},\delta_{A})= \left\{
\begin{array}{lll}
	 -\dfrac{k_{A}}{\delta_{A}}, & ~~~~~& \delta_{D} \leq d \\
	~ & ~ \\
   \dfrac{(\delta_{D} - d)^2}{2\delta_{D}\delta_{A}} - \dfrac{k_{A}}{\delta_{A}}, & ~~~~~~& d \leq \delta_{D} \leq d + \delta_{A} \\
   ~ & ~ \\
   1 - \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - \dfrac{k_{A}}{\delta_{A}} &~~~~~& \delta_{D} \geq d + \delta_{A}

\end{array}
\right.
\end{displaymath}

>>>>

\begin{figure}[hbtp]
\centering
\includegraphics[scale=1.4]{Images/functie4.pdf} 
\caption{The benefit function of the defender for a cost $k_{D} =1$, $d =2$ and $\delta_{A}=1.5$. The three pieces match with the three colours.}
\label{pfb}
\end{figure}

The benefit formula for the attacker, see figure \ref{pfa}:
\begin{displaymath}
\beta_{A}(\delta_{D},\delta_{A})= \left\{
\begin{array}{lll}
	 -\dfrac{k_{A}}{\delta_{A}}, & ~~~~~& \delta_{D} \leq d \\
	~ & ~ \\
   \dfrac{(\delta_{D} - d)^2}{2\delta_{D}\delta_{A}} - \dfrac{k_{A}}{\delta_{A}}, & ~~~~~~& d \leq \delta_{D} \leq d + \delta_{A} \\
   ~ & ~ \\
   1 - \dfrac{\delta_{A}}{2\delta_{D}} - \dfrac{d}{\delta_{D}} - \dfrac{k_{A}}{\delta_{A}} &~~~~~& \delta_{D} \geq d + \delta_{A} 
\end{array}
\right.
\end{displaymath}

\begin{figure}[hbtp]
\centering
\includegraphics[scale=1.4]{Images/benfa.pdf} 
\caption{The benefit function of the attacker for a cost $k_{A} =1.2$, $d =0.5$ and $\delta_{D}=5$. The three pieces match with the three colours.}
\label{pfa}
\end{figure}

### 2015-08-18 doc/template/conclusion.tex
shizzle

\chapter{Conclusion}
\label{cha:conclusion}
The final chapter contains the overall conclusion. It also contains
suggestions for future work and industrial applications.

\section{trala}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:

>>>>

\chapter{Conclusion}
\label{chapter5:conclusion}

This paper presents an adaptation to the basic FlipIt game by \citep{FlipIt} to model an attacker with a delay.

\section{General results and conclusions}
Our FlipIt model with worm propagation delay showed us some interesting results. 
\begin{itemize}
\item When the defender plays faster than the delay, the attacker will either have a negative benefit or a benefit equal to zero if the flips do not cost anything. It is for the defender a target to be able to play at a rate smaller or equal to the delay.
\item If the defender can play with a cost equal to zero, the attacker will not play. The defender can play at any rate he wants, and the attacker is always disadvantaged by his delay.
\item If the cost of the defender is non-zero, then, depending on the ratio between it's cost and the delay, from a certain value for the speed of the attacker, the defender will not play. The same thing goes for the attacker. This result is similar to the results obtained for the original FlipIt game.
\end{itemize}

While the first two results are intuitively clear, in the third case, the mathematical analysis is important to determine the exact ratio's at which it is no longer beneficial for the defender to play. \\
In the next section we list a set of extensions that might be interesting for further research.
\section{Further work}

\begin{description}
\item \textit{Nash equilibria}\\ This paper concluded with the best responds functions of the attacker and the defender. The next step would be to calculate the Nash equilibria. For these calculation multiple cases have to be examined, i.e. all the possible cases regarding the mutual relations between $k_{A}, k_{D}$ and $d$. 
\item \textit{Analysing other renewal strategies} \\ By analysing other renewal strategies, it might be interesting to find out if the original result regarding periodic versus other renewal strategies of the basic FlipIt game still stands. This result implied that periodic strategies strongly dominate the other renewal strategies if an opponent plays with a non-adaptive strategy. 
\item \textit{Delay for the defender}\\ This paper only assumed that the attacker had a delay. It could be interesting to have a defender with a delay. This situation could happen if it takes some time to make a patch to a system or when a new exploit is found. 
\item \textit{Variable delay} \\This paper assumed that the propagation delay has a fixed value. This can be relaxed by giving the attacker a delay could vary. The delay always negatively impacts the benefit of the attacker, so the attacker will always choose the lowest delay. But if the attacker always sends a new worm at every flip, it could be that the delays vary. In that regard, a variable delay would be more accurate to simulate a real world case. 
\item \textit{The defender flips a subset of nodes} \\ Our model assumes that if the defender flips, he flips the whole network. An option for the defender would be to flip a subset of nodes in the network. If the costs of flipping are related to the amount of nodes that are flipped, this option might be interesting to look at. It could be that a certain subset of nodes is found that protects the whole network. The PageRank matrix explained in Chapter 5 can help finding the right nodes. \\

\end{description}

As of today APTs become more and more extraordinary pieces of malicious code. There are APTs known to survive military-grade disk wiping and reformatting, and even after reformatting and reinstalling the operating system they are able to send sensitive data (Equation group \citep{Equation}). This means that even if you know that an APT is on your network, the actual practical ``flip'' has to exist and be known by the defender.  \\
Most often the best way to secure a system is to find the weakest link. This is mostly the employee. An effective way to counter this problem is to raise awareness about risks of infections amongst the employees.  Even in the presence of good security protection layers, spam filters, and firewalls, the opening of a phishing mail or use of an unknown USB stick by one of your employees may be suficient to infect a PC on the network, from which the virus can spread further. A system can however never be 100\% waterproof.  If an infection has occurred and the defender knows a countermeasure, the FlipIt game will help to determine the most effective pace at which countermeasures need to be taken.
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End:

### 2015-08-18 doc/template/nash.tex
shizzle

\section{Best responses of $\delta_{D}$}
The optimum functions will be piecewise functions. We distinguish six cases for different values of $\delta_{A}$ depending on the values of $k_{D}$ and $d$. 
%We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates.

@@

Combining those two leads to a rate for the defender equal to $[0,d]$. This means that the defender will never not play, which is intuitively correct if we know that the defender has no disadvantage to play because there is no cost involved.
\subsection*{$\delta_{A}=0$}
If $\delta_{A}=0$ it means that the attacker is playing as fast as possible. The defender has still a cost for every flip.
\begin{itemize}
\item if $k_{D} > d$, we look at the first piece of $opt_{D}(\delta_{A})$. It follows that the defender will not play (rate equal to $\infty$).
\item if $k_{D}=d$, corresponds to the second piece: the defender will play with a rate of $[d,\infty]$.
\item if $k_{D} < 0$, for the last case the rate of the defender corresponds to \textit{d}.
\end{itemize}
~~\\
   $  \infty \& \delta_{A}=0 ~~ \& ~~ k_{D} > d$ \\
Combining the results leads to a rate for the defender equal to $[d,\infty]$. 
\subsection*{$k_{D}=0$}
The cost of flipping for the defender is equal to zero. If we look at the benefit function of the defender and it's derivative in figure \ref{cost0}, we can see that the benefit for case B increases until $d$ and that the benefit in case A stays the same until $d$. After point $d$ the benefit decreases in all the cases.

@@

\begin{figure}[hbtp]
\centering
\includegraphics[scale=1]{Images/cost1.pdf} 
\caption{Benefit function and derivative for a cost $k_{D} > 0$ and with $d (=2)$ and $\delta_{D}=(1.5)$ not equal to zero.}
\label{cost1}
\end{figure}

@@

\section{Conclusion}
The Nash Equilibria can be found as the intersection points of the piecewise functions $opt_{A}(\delta_{D})$ and $opt_{D}(\delta_{A})$. For this we have to compare the function in terms of the relationship of the players move cost. We distinguish three cases: $k_{A} < k_{D} , ~k_{A} > k_{D} $ and $k_{A} = k_{D}$.

>>>>

\section{Best responses of $\delta_{D}$}
The optimum functions will be piecewise functions. We distinguish six cases for different values of $\delta_{A}$ depending on the values of $k_{D}$ and $d$. 
%We point out that $\delta_{A}$ and $\delta_{D}$ are positive rates. 
Figure XXX shows on the left how the three cases can be combined in a single continuous function (shown as the thick line). On the right, the derivative function indicates at which points the benefit will change. It is interesting to note that the derivative is also continuous, meaning that the benefit function is smooth and has no rough edges or corners. This implies that there are no sudden changes in benefit for small changes of period.

@@

Combining those two leads to a rate for the defender equal to $[0,d]$. This means that the defender will never not play, which is intuitively correct if we know that the defender has no disadvantage to play because there is no cost involved.
\subsection*{$\delta_{A}=0$}
If $\delta_{A}=0$ it means that the attacker is playing as fast as possible. The defender has still a cost for every flip. This means we have to look at different values for $k_{D}$ and $d$.
\begin{itemize}
\item if $k_{D} > d$, it follows that the defender will not play (rate equal to $\infty$). This is already the case for the first piece in the piecewise function. $\delta_{A}$ will always be smaller than $2(k_{D}-d)$ if $k_{D} > d.$
\item if $k_{D}=d$, it corresponds to the second piece: the defender will play with a rate of $[d,\infty]$.
\item if $k_{D} < 0$, will correspond with the last case, where the rate of the defender is equal to \textit{d}.
\end{itemize}
By combining the last two pieces we have  $d ~~for~~ \delta_{A} =0 ~~ and ~~k_{D} \leq d$.\\

\subsection*{$k_{D}=0$}
The cost of flipping for the defender is equal to zero. If we look at the benefit function of the defender and it's derivative in figure \ref{cost0}, we can see that the benefit for case B increases until $d$ and that the benefit in case A stays the same until $d$. After point $d$ the benefit decreases in all the cases.

@@

@@

\section{Conclusion}

The piecewise functions $opt_{D}$ and $opt_{A}$ give some interesting results. \todo{resultaten nog toevoegen, zelfde als in de andere conclusie}

The Nash Equilibria can be found as the intersection points of the piecewise functions $opt_{A}(\delta_{D})$ and $opt_{D}(\delta_{A})$. For this we have to compare the function in terms of the relationship of the players move cost. We distinguish three cases: $k_{A} < k_{D} , ~k_{A} > k_{D} $ and $k_{A} = k_{D}$.
